\section*{\hspace{17mm}Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}
Both the development of the FjordOs model and the evaluation of the model's performance is part of the FjordOs project. FjordOs is a cooperation between MET Norway (Norwegian Meteorological institute), HSN (University College of Southeast Norway), NIVA (Norwegian Institute for Water Research), Kystverket (Norwegian Coastal Administration), ExxonMobil, FFI (Norwegian Defence Research Establishment), the counties Vestfold, Buskerud, and {\O}stfold, and AGNES AB Milj{\o}konsulent. It is funded in parts by the Regional Research Fund Oslofjordfondet through Research grant no. 226022, and through smaller contributions from MET Norway, NIVA, HSN, Kystverket, and ExxonMobil. We gratefully acknowledge the support. 

We also like to thank the crew and captain of the R/V Trygve Braarud for excellent support during the cruises and for enduring all our requests. Moreover we would like to thank ExxonMobil for making their acoustic Doppler measurements outside of Slagentangen available to us. Likewise we thank NIVA who provided the CTD data, Norwegian Mapping Authority for available observations of water level, Scanmar AS for temperature measurements near {\AA}sg{\aa}rdstrand, and Finnerud Elektronikk for making the beach temperatures at Stor{\o}yodden, Hvalstrand, and Sj{\o}strand available to us. Finally we thank NIVA and Statnett for providing the data from the four acoustic Doppler profiling measurements across the fjord from Sm{\aa}skj{\ae}r to Evje, and from the two stations at Filtvedt and Brenntangen. 




